---
title: "Analysis of t-cell populations"
author: "Kent Riemondy RBI"
date: "`r Sys.Date()`"
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "html") })
output:
  html_document:
    theme: yeti
    toc: true
    toc_float: true
    toc_depth: 3
    highlight: kate
    fig_caption: true
    code_folding: hide
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```


```{r, message = FALSE}
# source(here::here("src/utils.R"))
source(here::here(file.path("..", "..", "..", "src/utils.R")))

seed_value <- 20210805
library(ggbeeswarm)
library(ggpubr)
library(rstatix)
```

```{r}
so <- qread(file.path("objects", "so_lym.qs"), nthreads = 4)
```

```{r}
is_old_mdata <- all(so@meta.data[so$UPN == "1455",c("Dx")] == "low grade glioma FGFR mutant")
if(is_old_mdata){
  so@meta.data[so$UPN == "1455", "Dx"] <- "ganglioglioma"
  so@meta.data[so$UPN == "1455", "Dx_simple"] <-"ganglioglioma"
  so@meta.data[so$UPN == "1171", "Dx"] <- "pilocytic astrocytoma"
  so@meta.data[so$UPN == "1171", "Dx_simple"] <-"pilocytic astrocytoma"
}
```

## Overview 

### UPN

The clustering is somewhat driven by each patient specific sample, so  batch correction will be necessary. We may also need to pursue a different approach based on clustering each patient sample individually then merging related profiles. 

```{r}
plot_umap(so, "UPN")
```
```{r}
plot_umap(so, "t_res.0.3", legend_title = "clusters") 
```

### UPN after correction

The clustering is now less strongly driven by each patient specific sample

```{r}
plot_harmony(so, "UPN")
```
```{r}
plot_harmony(so, "t_harmony_res.0.3", legend_title = "harmony-clusters") 
```

## Compare cluster markers to immune related cell atlas {.tabset}

I've been collating public single cell and related datasets into a resource to use as a reference to compare cell types across datasets. By comparing to an atlas we will be able to annotate cell types in the experiment with high confidence. 

The reference datasets that I will use for this analysis include:

Immune  cell types from our recent medulloblastoma paper in [NeuroOncology](https://academic.oup.com/neuro-oncology/advance-article/doi/10.1093/neuonc/noab135/6291352) . These cell types will have a suffix `MB` or `MB-immune`.  

Immune  cell types from recent Ependymoma in [Cell Reports](https://www.sciencedirect.com/science/article/pii/S2211124720310081?via%3Dihub) . These celltypes will have a suffix of `EPN`  

Lastly a dataset from PBMC will be included to identify lymphocytes, with a suffix of `pbmc`.  


### Refined clusters

First the refined clustering from the unaligned data will be compared. 

```{r, fig.width=9, fig.height=9, message=FALSE, warning=FALSE, fig.cap="Heatmap showing the correlation coefficients of each cluster against various cell types from a diverse set of data."}
library(clustifyr)
library(clustifyrdatahub)
library(canatlas)

atlas <- list(
  MB_immune = canatlas::mb_atlas$immune,
  EPN_immune = canatlas::epn_atlas$immune,
  pbmc = canatlas::hpbmc[[2]]
)

mb_ids <- tibble(short_id = str_match(colnames(canatlas::mb_atlas$immune),
                            "\\((.+)\\)")[, 2], 
       long_id = str_remove(colnames(canatlas::mb_atlas$immune), 
                            " \\(.+\\)"), 
       original_id = colnames(canatlas::mb_atlas$immune),
       atlas_id = str_c(short_id))
colnames(atlas[["MB_immune"]]) <- mb_ids$atlas_id
atlas <- imap(atlas, 
              ~{colnames(.x) <- str_replace_all(colnames(.x), " ", "-") %>%
    str_remove("[0-9]{2}-") %>% 
    str_replace_all("-", "_") %>% 
    str_c(".", .y)
              .x})

to_keep <- Reduce(intersect, map(atlas, rownames)) 
atlas <- map(atlas, ~.x[to_keep, ]) %>% unname() %>% do.call(cbind, .)

res <- clustify(so,
                ref_mat = atlas,
                cluster_col = "t_harmony_res.0.3", 
                obj_out = FALSE)

hmap_mat <- t(res) %>% 
  .[rowSums(. > 0.6) > 0, ]

hmap <- Heatmap(hmap_mat,
        name = "Spearman\ncorrelation",
        row_title = "Reference cell types",
        row_names_gp = gpar(fontsize = 6),
        column_title = "Clusters",
        column_title_side = "bottom",
        col = viridis::magma(256))

# pdf(file.path(fig_dir, "clustifyr_cell_type_annotation_heatmap_refined_clusters.pdf"),
#     width = 9,
#     height = 16)
draw(hmap)
# null <- dev.off()

# hmap
```

<br>

<br>

Shown below are the top 5 most correlated cell types for each cluster. The correlation values are shown in parentheses. 

```{r}
top_calls <- rownames_to_column(as.data.frame(res), "cluster") %>% 
  pivot_longer(cols = -cluster) %>% 
  group_by(cluster) %>% 
  arrange(desc(value)) %>% 
  dplyr::slice(1:5) %>% 
  ungroup() %>% 
  mutate(cluster = factor(cluster, levels = as.character(sort(as.numeric(unique(cluster))))))  %>% 
  arrange(cluster) %>% 
  group_by(cluster) %>% 
  mutate(called_cell_type = str_c(name, " (", signif(value, 3), ")"),
         row_id = row_number()) %>% 
  select(cluster, 
         called_cell_type,
         row_id) %>% 
  pivot_wider(names_from = cluster,
              values_from = called_cell_type) %>% 
  select(-row_id)
top_calls
```
Shown below is the top  most correlated cell type for each cluster. Any reference celltype with a correlation < 0.6 will be annotated as unassigned. 

```{r}
top_call <- cor_to_call(res, threshold = 0.6) %>%  
  select(cluster, 
         called_cell_type = type, 
         correlation_value = r)
top_call
```


```{r}
if(!"t_cell_type" %in% colnames(so@meta.data)){
  # add subcluster ids
top_call_w_subclusters <- top_call %>%
  group_by(called_cell_type) %>% 
  mutate(n = n(),
         simple_id = str_split(called_cell_type, fixed(".")) %>% 
                map_chr(1), 
         t_cell_type_subcluster = ifelse(
           n == 1,
           simple_id,
           str_c(simple_id, "(", row_number(), ")"))) %>%
  select(cluster,
         t_cell_type = called_cell_type, 
         t_cell_type_subcluster)

so <- left_join(get_metadata(so), 
          top_call_w_subclusters,
          by = c("t_harmony_res.0.3" = "cluster")) %>% 
 select(cell, t_cell_type, t_cell_type_subcluster) %>% 
  column_to_rownames("cell") %>% 
  AddMetaData(so, .)
}


plot_harmony(so, 
          c("t_cell_type", "t_cell_type_subcluster"), 
          legend_title = "",
          sorted = "random") 
```

## Heatmap with annotated cell types

Same heatmap as above but showing the annotated cell type rather than the cluster #s

```{r, fig.width=9, fig.height=9,}
res <- clustify(so,
                ref_mat = atlas,
                cluster_col = "t_cell_type_subcluster", 
                obj_out = FALSE)

hmap_mat <- t(res) %>% 
  .[rowSums(. > 0.6) > 0, ]

hmap <- Heatmap(hmap_mat,
        name = "Spearman\ncorrelation",
        row_title = "Reference cell types",
        row_names_gp = gpar(fontsize = 6),
        column_title = "Called cell type",
        column_title_side = "bottom",
        col = viridis::magma(256))

# pdf(file.path(fig_dir, "clustifyr_cell_type_annotation_heatmap_refined_clusters.pdf"),
#     width = 9,
#     height = 16)
draw(hmap)
```
  
## Get markers 

Markers are saved to excel spreadsheets.

```{r}
library(scran)
so$Dx_simple <- ifelse(str_detect(so$Dx, "low grade glioma"),
                       "low grade glioma",
                       so$Dx)
# temporary fix for issue #4639
so_diet <- DietSeurat(so, graphs = "pca")
sce <- as.SingleCellExperiment(so_diet)
rm(so_diet)

markers.cell_type.up <- findMarkers(sce, 
                                groups = sce$t_cell_type, 
                                block = sce$orig.ident,
                                lfc = 0.5,
                                direction="up") %>% 
  lapply(., as.data.frame) %>%
  map(~rownames_to_column(.x, "gene") %>% 
        filter(FDR < 0.01))

markers.cell_type.up %>% 
  bind_rows(.id = "cluster") %>% 
  write_tsv(file.path(mkrs_dir, "scran_t_cell_type_consensus_markers.tsv"))

map(markers.cell_type.up, 
    ~set_xlsx_class(.x, "gene", "Text")) %>% 
openxlsx::write.xlsx(., file.path(mkrs_dir, "scran_t_cell_type_consensus_markers.xlsx"),
                     overwrite = TRUE)

markers.cell_type_sub.up <- findMarkers(sce, 
                                groups = sce$t_cell_type_subcluster, 
                                block = sce$orig.ident,
                                lfc = 0.5,
                                direction="up") %>% 
  lapply(., as.data.frame) %>%
  map(~rownames_to_column(.x, "gene") %>% 
        filter(FDR < 0.01))


markers.cell_type_sub.up %>% 
  bind_rows(.id = "cluster") %>% 
  write_tsv(file.path(mkrs_dir, "scran_t_cell_type_subcluster_consensus_markers.tsv"))

map(markers.cell_type_sub.up, 
    ~set_xlsx_class(.x, "gene", "Text")) %>% 
openxlsx::write.xlsx(., file.path(mkrs_dir, "scran_t_cell_type_subcluster_consensus_markers.xlsx"),
                     overwrite = TRUE)
```


```{r}
mkrs <- get_marker_summaries(so, 
                     "t_cell_type", 
                     outdir = mkrs_dir, 
                     min_pct = 10,
                     prefix = "t_cell_type")
```

```{r}
mkrs <- get_marker_summaries(so, 
                     "t_cell_type_subcluster", 
                     outdir = mkrs_dir, 
                     min_pct = 10,
                     prefix = "t_cell_type_subcluster")
```


## Cell counts and changes in cell proportions between tumor types

```{r, fig.width = 7}
plot_harmony(so, 
             "t_cell_type_subcluster",
             group = "Dx_simple",
             legend_title = "") 
```

```{r, rows.print = 50}
cell_count_mat <- get_cell_count_matrix(so, "UPN", "t_cell_type_subcluster") %>% 
  rownames_to_column("tumor") 

write_csv(cell_count_mat, file.path(tbls_dir, "t_cell_type_subcluster_cell_counts.csv"))
cell_count_mat
```

```{r}
per_patient <- group_by(get_metadata(so), UPN) %>%
      mutate(n_cells = n()) %>%
      group_by(UPN,
               t_cell_type_subcluster,
               Dx_simple) %>%
      summarize(n = n(),
                prop_cell_type = n / unique(n_cells))

stat.test <- per_patient %>%
  group_by(t_cell_type_subcluster) %>%
  wilcox_test(prop_cell_type ~ Dx_simple) %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_significance("p.adj") %>% 
  add_xy_position(x = "t_cell_type_subcluster", dodge = 1)

p <- ggplot(per_patient, aes(t_cell_type_subcluster, prop_cell_type)) +
  geom_boxplot(aes(fill = Dx_simple),
               size = 1,
               alpha = 0.6,
               coef = 1e5) +
  geom_beeswarm(aes(color = Dx_simple),
                dodge.width=1,
                size = 0.75)  +
  stat_pvalue_manual(stat.test,  
                     label = "{p.adj}{p.adj.signif}", 
                     tip.length = 0, 
                     hide.ns = T) +
  scale_fill_manual(values = palette_OkabeIto) + 
  scale_color_manual(values = palette_OkabeIto) +
  labs(x ="",
       y = "Proportion of cells",
       color = "",
       fill = "") +
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.5))

p
save_plot(file.path(fig_dir, "t-cell-subpopulations-per-dx.pdf"),
          p,
          base_asp = 2)
```

Statistics for pairwise comparisions using wilcox rank sum test, corrected using bonferroni. Using a count based test (e.g. edgeR) may be more sensitive

```{r, rows.print=50, cols.min.print=5}
stat.test  %>% 
  select(1, group1, group2, p:p.adj.signif, n1:statistic)
```

```{r}
per_patient <- group_by(get_metadata(so), UPN) %>%
      mutate(n_cells = n()) %>%
      group_by(UPN,
               t_cell_type_subcluster,
               mutation_status) %>%
      summarize(n = n(),
                prop_cell_type = n / unique(n_cells))


stat.test <- per_patient %>%
  group_by(t_cell_type_subcluster) %>%
  wilcox_test(prop_cell_type ~ mutation_status) %>%
  adjust_pvalue(method = "bonferroni") %>%
  add_significance("p.adj") %>% 
  add_xy_position(x = "t_cell_type_subcluster", dodge = 1)

p <- ggplot(per_patient, aes(t_cell_type_subcluster, prop_cell_type)) +
  geom_boxplot(aes(fill = mutation_status),
               size = 1,
               alpha = 0.6,
               coef = 1e5) +
  geom_beeswarm(aes(color = mutation_status),
                dodge.width=1,
                size = 0.75)  +
  stat_pvalue_manual(stat.test,  
                     label = "{p.adj}{p.adj.signif}", 
                     tip.length = 0, 
                     hide.ns = T) +
  scale_fill_manual(values = palette_OkabeIto) + 
  scale_color_manual(values = palette_OkabeIto) +
  labs(x ="",
       y = "Proportion of cells",
       color = "",
       fill = "") +
  theme(axis.text.x = element_text(angle = 90,
                                   hjust = 1,
                                   vjust = 0.5))

p
save_plot(file.path(fig_dir, "t-cell-subpopulations-per-mutation-type.pdf"),
          p,
          base_asp = 2)
```

Statistics for pairwise comparisions using wilcox rank sum test, corrected using bonferroni. Using a count based test (e.g. edgeR) may be more sensitive

```{r, rows.print=50, cols.min.print=5}
stat.test  %>% 
  select(1, group1, group2, p:p.adj.signif, n1:statistic)
```

```{r}
plot_cell_proportions(so, 
                      sample_id = "UPN",
                      group_id = "t_cell_type_subcluster",
                      facet_by = "Dx_simple"
                      ) +
  labs(x = "UPN",
       fill = "")
```
## Go Terms

```{r}
library(gprofiler2)
library(ggrepel)
# Use top 200 markers 
mkrs <- read_tsv(file.path("markers", 
                           "t_cell_type_subcluster_cluster_markers.tsv")) %>% 
  filter(logFC > 0,
         pct_in > 10,
         padj < 0.05) %>% 
  group_by(group) %>% 
  dplyr::slice(1:200)
mkrs_split <- split(mkrs, mkrs$group) %>% 
  map(~pull(.x, feature))
go_res <- gost(mkrs_split, 
               ordered_query = TRUE,
               sources = c("GO", 
                          "KEGG", 
                          "REAC", 
                          "TF"))
go_res[["result"]] %>% 
  mutate(parents = unlist(map(parents, ~str_c(.x, collapse = ",")))) %>% 
  write_tsv(., file.path(tbls_dir, "goterms_t_cell_type_subclusters.tsv"))
go_res[["result"]] %>% 
  mutate(parents = unlist(map(parents, ~str_c(.x, collapse = ","))),
         source = str_replace_all(source, "[[:punct:]]", " ")) %>% 
  select(-significant) %>% 
  split(., .$query) %>% 
  openxlsx::write.xlsx(., 
                       file.path(tbls_dir, "goterms_t_cell_type_subclusters.xlsx"),
                       overwrite = TRUE)
```

```{r}
go_res_all <- gost(mkrs_split, 
               ordered_query = TRUE, 
               significant = FALSE,
               sources = c("GO", 
                          "KEGG", 
                          "REAC", 
                          "TF"))
go_res_to_plot <- go_res_all[["result"]] %>% 
  mutate(parents = unlist(map(parents, ~str_c(.x, collapse = ","))),
         source = str_replace_all(source, "[[:punct:]]", " ")) %>% 
  select(-significant) 
```

```{r, fig.height = 9, fig.width = 9}
dbs <- c("GO BP", "GO MF", "GO CC")
top_n_terms <- 10
gterm_hmaps <- map(dbs,
    ~{
  res <- filter(go_res_to_plot, 
                   source == .x,
                   p_value < 0.05) %>% 
    group_by(query) %>% 
    arrange(p_value, .by_group = TRUE) %>% 
    dplyr::slice(1:top_n_terms)
  
  to_plot <- filter(go_res_to_plot,
                    source == .x,
                    term_name %in% res$term_name) %>%
    select(query, p_value, term_name) %>% 
    pivot_wider(names_from = query, values_from = p_value) %>% 
    column_to_rownames("term_name") %>% 
    as.matrix() 
  
  to_plot[is.na(to_plot)] <- 1
  to_plot <- -log10(to_plot)
  hmap <- Heatmap(to_plot, 
          name = "-log10(pvalue)",
          column_title = .x,
          column_order = unique(res$query),
          row_order = unique(res$term_name),
          col = viridis::magma(256),
          cluster_columns = TRUE, 
          cluster_rows = FALSE, 
          row_names_side = "left",
          row_names_gp = gpar(fontsize = 8))
  hmap
})
walk(gterm_hmaps, draw)
```
 
## Overview umap figs

```{r}
reid <- c("CD8_TH1" = "CD8_TH1",
          "gdT" = "gdT",
          "NK_cells"= "NK",
          "NKT" = "NKT",
          "CD4" = "CD4",
          "CD4_TCM"= "CD4",
          "T(1)" = "CD8_dying",
          "T(2)" = "CD8",
          "Treg" = "Treg")
  
so$simple_t_cell_type <- reid[so$t_cell_type_subcluster]
p <- plot_harmony(so, 
                  "simple_t_cell_type", 
                  .cols = palette_OkabeIto, 
                  sorted = "random",
                  legend_title = "")

p

save_plot(file.path(fig_dir, "umap_by_simplified_t_cell_type.pdf"),
          p,
          base_asp = 1.4)
```


## Clustrees

```{r}
library(clustree)
p <- clustree(so, prefix = "t_harmony_res.") +
  scale_color_manual(values = discrete_palette_default)
p
save_plot(file.path(fig_dir, "t_harmony_clustree_0.3_selected.pdf"), 
          p,
          base_asp = 1.5,
          base_height = 8)
```

## Save data

```{r, eval = TRUE}
qsave(so, file.path("objects", "so_lym.qs"))
#so <- qread( file.path("objects", "so_lym.qs"))
```

## Session info
```{r}
sessioninfo::session_info()
```

